---
title: "STAT380 HW1"
author: Mia Iceland
output: html_document
date: "2024-01-14"
---

## Front Matter
```{r}
library(tidyverse)
```

## Problem 1
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length)) +
  geom_histogram(fill = "red", color = "black", binwidth = .25) +
  labs(x = "Sepal Length (in cm)",
       y = "Number of Flowers")
```

Flowers with the sepal length of 5 cm has the greatest population. As the Sepal length decreases and increases from 5 cm, the number of flowers decreases.

## Problem 2
```{r}
ggplot(data = iris, mapping = aes(x = Species)) +
  geom_bar(fill = "white", color = "black") +
  labs(x = "Species", y = "Count")
```

This graph proves that the Iris study provided the same amount of cases for each species of flower: 50 Setosa, Versicolor, and Viginica flowers for a total of 150 cases recorded. 

## Problem 3
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point(shape = 6) +
  labs(x = "Sepal Length (in cm)",
       y = "Sepal Width (in cm)") +
  geom_smooth(method = lm, se = FALSE)
```

What I have learned from this graph is that as Sepal Length increases Sepal Width slightly decreases. There is a slightly negative, linear relationship between Sepal Length and Sepal Width.

## Problem 4
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  labs(x = "Sepal Length (in cm)",
       y = "Sepal Width (in cm)") +
  geom_smooth(method = lm, se = FALSE, aes(group = Species))
```

When the flowers are grouped by species, the graph shows that there is a positive, linear relationship between sepal length and sepal width for all three species. 

## Problem 5
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("orange", "darkblue", "green")) +
  labs(x = "Sepal Length (in cm)",
       y = "Sepal Width (in cm)") +
  geom_smooth(method = lm, se = FALSE, aes(group = Species))
```

## Problem 6
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  scale_color_manual(values = c("orange", "darkblue", "green")) +
  labs(x = "Sepal Length (in cm)",
       y = "Sepal Width (in cm)") +
  geom_smooth(method = lm, se = FALSE, aes(group = Species))
```

"how to change color of points in ggplot"

<https://stackoverflow.com/questions/43770579/how-to-change-the-color-in-geom-point-or-lines-in-ggplot>

## Problem 7
```{r}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  labs(x = "Sepal Length (in cm)",
       y = "Sepal Width (in cm)") +
  facet_grid(~Species) +
  theme_bw() +
  geom_smooth(method = lm, aes(group = Species))
```

"how to remove the gray background from a ggplot graph"

<https://www.geeksforgeeks.org/remove-grid-and-background-from-plot-using-ggplot2-in-r/>

## Problem 8
```{r}
ggplot(data = iris, mapping = aes(x = Species, y = Sepal.Width)) +
  geom_boxplot() +
  labs(x = "Species",
       y = "Sepal Width (in cm)") +
  theme_bw()
```


